---
title: "hw2_hs_AirbnbNYC"
author: "Huan Sun"
date: "3/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing the Dataset & Packages
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(maps) # The maps package has geographic information on all U.S states
library(ggmap)
library(tmap)
library(rgdal)
library(rgeos)
library(geojsonio)
library(viridis)
library(hrbrthemes)
library(tm)
library(sp)
library(DT)
library(devtools)
library(leaflet)
library(ggthemes)
library(RColorBrewer)
```

```{r}
# use getwd() and setwd() to change view and change working directory
data<-read.csv("data/airbnb_listings/airbnb_listings.csv")

airbnbnyc <- data%>%
  select('id','host_id','host_listings_count','latitude','longitude','room_type','accommodates','bathrooms','bedrooms','price','neighbourhood_cleansed','neighbourhood_group_cleansed','availability_365','number_of_reviews','review_scores_rating','transit')

head(airbnbnyc)
```

# 1. Overall Location

## a) Provide a map to show where in New York City AirBnB listings are located.

Read the `neighbourhoods.geojson` of NYC neighborhoods file, and `nybb` of NYC boroughs files in.

```{r}
# read the initial shape file with neighbourhoods and borough data
nyc_neighborhoods <- readOGR(dsn="data/neighbourhoods.geojson")
# convert the shape object so that it may be portrayed on a map 
nyc_neighborhoods <- spTransform(nyc_neighborhoods,CRS("+proj=longlat +datum=WGS84"))
nyc_neighborhoods<-fortify(nyc_neighborhoods)

nyc_boroughs <- readOGR("data/nyc_boroughs_map/.","nybb")
nyc_boroughs<- spTransform(nyc_boroughs,CRS("+proj=longlat +datum=WGS84"))
nyc_boroughs<- fortify(nyc_boroughs)
```

```{r}
# get nyc background map in with get_map
map_nyc <- get_map("New York City",zoom = 11,source = "stamen",maptype = "toner-lite")
```

```{r}
# plot the map with ggmap, which returns a ggplot object
g <- ggmap(map_nyc,
           base_layer=ggplot(aes(x=long,y=lat),
                             legend=FALSE,
                             data=nyc_boroughs),
           extent = "normal", maprange=FALSE)+ 
  geom_polygon(aes(x=long, y=lat, group=group, fill=id,alpha=3), 
               size=0.5, color='#636363',alpha=0.3)+
  geom_point(data=airbnbnyc,aes(x=longitude,y=latitude),size=0.3, alpha=0.1, color="#2c7fb8")+
  theme_map()+
  theme(legend.position = "none")
g 
```
According to the map, the highly density area are mostly in Manhattan and Brooklyn, but we'll get a closer look of the map with the density function in the below question.

## b) Provide a map in which you _summarize the density_ of the AirBnB listings and _highlight the hot-spots_ for AirBnB locations. Make sure to _annotate_ a few hot-spots on the map.
```{r}
# have a closer look at metropolitan area 
map_nyc <- get_map("New York City",zoom = 12,source = "stamen",maptype = "toner-lite")
```

 
```{r}
# highlight the density of hot airbnb spots in nyc
density<-ggmap(map_nyc)+
  stat_density2d(data = airbnbnyc, geom = "polygon",
  aes(x = longitude, y = latitude, fill=..level.., alpha=..level..)) + 
  scale_fill_distiller(palette=4, direction=1) +
  theme(legend.position="bottom")+
  theme_map()+
  ggplot2::annotate("text",x=-73.994508, y=40.763186, label="Hells Kitchen",
                   color="Dark Blue",fontface=4, size=3) +
  ggplot2::annotate("text",x=-73.98736, y=40.72527, label="East Village",
             color="Dark Blue",fontface=4, size=3)+
  ggplot2::annotate("text",x=-73.951996, y=40.71200, label="Williamsburg",
             color="Dark Blue",fontface=4, size=3)+
  theme(legend.position = "none")
  

density

```
 It is not surprising to find that the southwestern part of Central Park area and East Village area has most house rented as Airbnb, but also interesting to find that Willimsburg that as the uprising popular neighborhood had gained much attention.
 
# 2. Renting out your apartment vs. permanent rentals.

An Airbnb host can set up a calendar for their listing so that it is only available for a few days or weeks a year. Other listings are available all year round (except for when it is already booked). Entire homes or apartments highly available and rented frequently year-round to tourists probably don't have the owner present, are illegal, and more importantly, are displacing New Yorkers.

Hint: The variable `availability_365`: What part of the year is the property available to be rented is a possible choice to categorize rentals.

## a) Choose a combination of both _maps and non-mapping visualizations_ (graphs or tables) to explore *where* in NYC listings are available sporadically vs. year-round. Make sure to _highlight_ the neighborhoods were _most listings appear to be permanent_ or semi-permanent rentals.

#### - Graphic distribution of locations of year-round listings.

```{r}
# set a definition for each case
airbnbnyc_case <- airbnbnyc %>%
  mutate(case=case_when(airbnbnyc$availability_365 >= 270 ~ "Permanent",
            airbnbnyc$availability_365 >= 180 ~ "Semi-Permanent",
            airbnbnyc$availability_365 < 180 ~ "Sporadically"))

# check which neighborhood has the most permanent rental
permanent_count <- airbnbnyc_case %>%
  group_by(neighbourhood_cleansed,case)%>%
  summarize(nbh_number=n())%>%
  mutate(group_sum=sum(nbh_number),by=neighbourhood_cleansed)%>%
  arrange(desc(group_sum))%>%
  rename(neighborhood=neighbourhood_cleansed,
         neighbor_case_total=nbh_number,
         neighbor_total=group_sum)%>%
  select(-by)
  
head(permanent_count)
```
```{r}
# create a table for presentation
datatable(permanent_count, 
          rownames=FALSE,colnames=c("Neighborhood","Case","Case Number in Neighborhood","Total Number in Neighborhood"),
          caption=htmltools::tags$caption("Listing Availability: Sporadically or Year-Round"), 
          #options=list(autoWidth = TRUE, dom = "ft", pageLength=10), 
          filter = list(position="top"))
```

```{r}
# select the top 10 neighborhoods with most listings
permanent_count_top10 <- permanent_count%>%
  head(30)

# make a stacked bar chart for comparision
ggplot(data=permanent_count_top10,
       aes(x=reorder(neighborhood,neighbor_total),y=neighbor_total,
           fill=case))+
geom_bar(stat="identity")+
  theme_classic()+
  labs(x="Number of Listing", y='Neighborhood', title='Listing Availability: Sporadically or Year-Round')+
  theme(plot.title=element_text(hjust=0.5))+
  coord_flip()+
  theme_tufte(base_size = 13) +
  scale_fill_manual(values=c('#ece2f0','#a6bddb','#1c9099'))
```

#### - Mapping for locations of year-round listings.

```{r}
g <- ggmap(map_nyc)+
  stat_density2d(data = airbnbnyc_case, geom = "polygon",
  aes(x = longitude, y = latitude, fill=..level.., alpha=..level..)) + 
  scale_fill_distiller(palette=4, direction=1) +
  theme_map()+
  facet_wrap(~case)+
  theme(legend.position="bottom")

g
```

```{r}
pal = colorFactor("Set2", domain = airbnbnyc_case$case)
color_case=pal(airbnbnyc_case$case)

leaflet(airbnbnyc)%>%
  addProviderTiles("Stamen.TonerLite") %>%
  addCircles(color=color_case)%>%
  addLegend(pal = pal, values = ~airbnbnyc_case$case, title = "Case")
```

## b) Some hosts (identified by `host_id`) operate multiple rentals. Provide a _data table of the the top hosts_, and the followings:
- total number of listings they are associated with
- the average nightly price
- the estimated average monthly total income from these listings

```{r}
# 1) calculate the number of listing each host has
host_listing_number<-airbnbnyc%>%
  group_by(host_id)%>%
  summarize(number=n())%>%
  arrange(desc(number))

# convert the price column in numeric types
airbnbnyc$price=gsub("\\$","",airbnbnyc$price)
airbnbnyc$price<-as.numeric((airbnbnyc$price))

# 2) calculate the average nightly price  
host_nightly_avg<- airbnbnyc%>%
  group_by(host_id)%>%
  summarize(nightly_avg=round(mean(price),2))%>%
  arrange(desc(nightly_avg))
```

```{r}
# join the above three newly created columns back onto the airbnbnyc file 
host_info<- airbnbnyc%>%
  left_join(host_listing_number,by=c("host_id"))%>%
  left_join(host_nightly_avg,by=c("host_id"))%>%
  select(host_id,availability_365,number,nightly_avg)%>%
  # here I think when calculating the estimate income, should only consider the days when the house is available to rent
  #3) calculate the estimate average monthly total income from the listings: should be the avg*(available_365)/12
  mutate(monthly_income=round(number*nightly_avg*availability_365/12),2)%>%
  select(host_id,number,nightly_avg,monthly_income)%>%
  group_by(host_id,number,nightly_avg)%>%
  mutate(monthly_income=round(mean(monthly_income),2))%>%
  arrange(desc(monthly_income))%>%
  distinct()

# create the datatable
datatable(host_info, 
          rownames=FALSE,colnames=c("Host Id","Listing Count","Nightly Avg Pirce","Monthly Avg Income"),
          caption=htmltools::tags$caption("Fun Facts about the Airbnb Hosts"), 
          #options=list(autoWidth = TRUE, dom = "ft", pageLength=10), 
          filter = list(position="top"))
```


# 3. Top Reviewed Rentals

## Provide an _interactive map_ which shows the _Top 100 most expensive_ and _Top 100 best reviewed rentals_ in NYC. 
The map should differentiate these two groups and upon clicking on a point on the map should show some basic information (at least 3 pieces of information) in a tool tip.

```{r}
# get the info of top100 most expensive and best review rentals
airbnbnyc_rank<-airbnbnyc%>%
  arrange(desc(price))%>%
  mutate(price_rank=row_number())%>%
  arrange(desc(review_scores_rating))%>%
  mutate(review_rank=row_number())

head(airbnbnyc_rank)
```

```{r}
leaflet(airbnbnyc_rank)%>%
  # prepare the base layers
  addTiles(group = "Toner") %>%
  addProviderTiles(providers$Stamen.TonerLite, group="Toner Lite") %>%
  # add first layer 
  addCircles(group="Top 100 Most Expensive", 
      data=subset(airbnbnyc_rank,airbnbnyc_rank$price_rank < 101),
      opacity = 1.0, stroke = TRUE,
      color = "#af8dc3", weight=1,
      popup = paste("AirBnb Id:",airbnbnyc_rank$id,"<br/>",  
        "Nightly Price:", airbnbnyc_rank$price, "$","<br/>", 
        "Accomodate:",  airbnbnyc_rank$accommodates, "People"))%>%
  # add second layer
  addCircles(group="Top 100 Best Reviews", 
      data=subset(airbnbnyc_rank,airbnbnyc_rank$review_rank < 101),
      opacity = 1.0, stroke = TRUE,
      color = "#7fbf7b", weight=1,
      popup = paste("AirBnb Id:",airbnbnyc_rank$id,"<br/>",  
        "Nightly Price:", airbnbnyc_rank$price, "$","<br/>", 
        "Accomodate:",  airbnbnyc_rank$accommodates, "People"))%>%
  # add layer control
  addLayersControl(
    baseGroups = c("OpenStreetMap","Toner Lite"),
    overlayGroups = c("Top 100 Most Expensive","Top 100 Best Reviews"),
    options = layersControlOptions(collapsed = TRUE) )
```